---
title: "CRIM 515 Project 01 Template"
author: "My Name"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---
# Libraries
```{r}
library(gtrendsR)
library(Hmisc)
library(rvest)
library(tidyverse)
```

# Research Question
All words, no code

# Literature Review
All words, no code.

# Data
Some words, and some code...

First, I acquired public interest data from ________.
```{r}
google.trends <- gtrends("mass shooting", geo = "US", time = "2014-01-01 2023-09-25")
```

Second, I used mass shooting data from _____.
```{r}

```

# Methods
Some words, and some code...
I created summary tables for ____ (events/victims), by ______ (year/month)...
```{r}

```
Some words about my break-point...
Some code prepping for my t-test...
```{r}

```

# Findings
## Finding 1: Mass Shooting Activity
Some words, some code (including the t test), a visual...
```{r}

```

## Finding 2: Public Interest about mass shootings
Some words, some code, and a visual...
```{r}

```

## Finding 3: Intersection of mass shootings and public interest
Some words, some code, and a visual...
```{r}

```

# So What
All words, no code.