---
title: "CRIM 515 Project 2 Template version 2"
author: "My Name"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---
## Libraries
```{r, message=FALSE}
library(gtrendsR)
library(tidyverse)
library(tidycensus)
library(rnaturalearth)
library(rnaturalearthdata)
library(maps)
library(lubridate)
library(usmap)
library(sf)
library(leaflet)
library(plotly)
theme_set(theme_bw())
```


## Research Question
*Note - I'd recommend doing this section at the end of the project*

Just text, 3 components...

1. how police shootings differ by victims and places over time;

2. if public interest about police shootings is changing; and 

3. describe the relationship between trends in events and public interest in police shootings.

## Literature Review
Just text, one citation, one paragraph

## Data
Some text about acquiring and cleaning three different datasets, and some code

### Dataset 1 (name it)
Add some sentences here about this data source - how many records, date range, etc...
```{r}
wapo.data <- read.csv("https://raw.githubusercontent.com/washingtonpost/data-police-shootings/master/v2/fatal-police-shootings-data.csv", 
                      stringsAsFactors = FALSE)
wapo.data$date <- as.Date(wapo.data$date)
wapo.data$month <- substr(wapo.data$date, 6, 7)
wapo.data$year <- substr(wapo.data$date,0,4)
wapo.data$yearmonth <- paste(wapo.data$year, wapo.data$month, sep = "-")
wapo.data$race <- gsub("B;H", "O", wapo.data$race)
```
##### Dataset 2 (name it)
Add some sentences here about this data source - how many records, date range, etc...
```{r}
today <- as.character(Sys.Date())
time.range <- paste("2015-01-01", today, sep = " ")
trends1 <- gtrends("police shooting", geo = "US", time = time.range)
```
##### Dataset 3 (name it)
Add some sentences here about this data source - how many records, date range, etc...

###### Table 1 - name it
```{r, message=FALSE}
race.2021 <- get_acs(geography = "state", 
                     variables = c("B02008_001", "B02009_001", "B02010_001", "B02011_001", "B03001_003"), 
                     year = 2021)

race.2021$variable <- gsub("B02008_001", "White", race.2021$variable)
race.2021$variable <- gsub("B02009_001", "Black", race.2021$variable)
race.2021$variable <- gsub("B02010_001", "Native American", race.2021$variable)
race.2021$variable <- gsub("B02011_001", "Asian", race.2021$variable)
race.2021$variable <- gsub("B03001_003", "Hispanic", race.2021$variable)
```
###### Table 2 - name it
```{r, message=FALSE}
race.total <- get_acs(geography = "us", 
                      variables = c("B02008_001", "B02009_001", "B02010_001", "B02011_001", "B03001_003"), 
                      year = 2021)

race.total$variable <- gsub("B02008_001", "White", race.total$variable)
race.total$variable <- gsub("B02009_001", "Black", race.total$variable)
race.total$variable <- gsub("B02010_001", "Native American", race.total$variable)
race.total$variable <- gsub("B02011_001", "Asian", race.total$variable)
race.total$variable <- gsub("B03001_003", "Hispanic", race.total$variable)
```
###### Table 3 - name it
```{r, message=FALSE}
totalpop.2021 <- get_acs(geography = "state", variables = "B01003_001", year = 2021)
names(totalpop.2021) <- c("GEOID", "State", "Variable", "Population", "junk")
totalpop.2021$State.PCT <- round(totalpop.2021$Population/sum(totalpop.2021$Population)*100,2)
temp.race.population <- race.2021 %>% left_join(totalpop.2021, by = "GEOID")
race.population <- temp.race.population[c(1:4,8,10)]
names(race.population) <- c("GEOID", "State", "Race", "Population.Race", "Population.State", "State.PCT")
race.population$Race.PCT <- round(race.population$Population.Race/race.population$Population.State*100, 2)
race.total$PCT <- race.total$estimate/sum(totalpop.2021$Population)*100
names(race.total) <- c("GEOID", "Area", "Race", "Count", "moe", "Race.PCT")
```

## Methods
Some text about stats and techniques, and some code. I do my work with summary tables here... and I only include the summary tables I actually analyze. And that analysis matches my research question. I would also do any event-census table joining here, too
```{r}

```

## Findings
#### Finding 1
Some analysis, and some code for Visual 1
```{r}

```

#### Finding 2
Some analysis, and some code for Visual 2
```{r}

```

#### Finding 3
Some analysis, and some code for Visual 3
```{r}

```

## So What
All text, two paragraphs (roughly), all text


# Example Section
```{r, echo=FALSE, message=FALSE, results='hide', fig.show='hide'}
# when you set 'echo' to false, you are hiding the block of code 
# when you set 'message' to false, you are hiding the console text that comes after running your code
# when you set 'results' to hide, you are hiding extra processing that occasionally shows up in the console
# when you set fig.show to hide, you are hiding the plot output
```
