---
title: "Course Name Course Number Project Template"
author: "Your Name"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}

```

# Research Question

Provide a brief paragraph (3-4 sentences) on the research question.

# Lit Review

Provide a paragraph (4-6 sentences) describing a prior study related to your current project.

# Data

Provide a brief paragraph (3-4 sentences) on the data used in this study. 

# Methods

Provide a brief paragraph (3-4 sentences) on the research methods leveraged to answer this question. 

# Findings
```{r}

```

### Finding 1

### Finding 2

### Finding 3

### Finding 4

### Finding 5

# Visuals

### Visual 1
```{r, echo=FALSE, message=FALSE, warning=FALSE}

```

### Visual 2
```{r, echo=FALSE, message=FALSE, warning=FALSE}

```

### Visual 3
```{r, echo=FALSE, message=FALSE, warning=FALSE}

```

# So What

Provide a brief paragraph (4-6 sentences) on the importance of this study, as a criminal justice professional. 
