---
title: "CRIM 515 Final Project In-Class"
author: "Your Name"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  cleanrmd::html_document_clean:
    theme: markdown-modest
    toc: yes
    toc_float: 
      collapsed: true
---

Some R Markdown styles to try, if you're interested:

“almond”, “awsm.css”, “axist”, “bamboo”, “bullframe”, “holiday”, “kacit”, “latex.css”, “markdown-splendor”, “markdown-retro”, “markdown-air”, “markdown-modest”, “marx”, “minicss”, “new.css”, “no-class”, “picocss”, “sakura”, “sakura-vader”, “semantic”, “simplecss”, “style-sans”, “style-serif”, “stylize”, “superstylin”, “tacit”, “vanilla”, “water”, “water-dark”, “writ”

# 1. Research Question

EXAMPLE:  Using 2008-2024 data, this project forecasts animal-related calls for service counts and locations in the City of Fairfax, VA for each month in 2025.

# 2. Literature Review

INSERT THAT TEXT, WITH CITATION, HERE.

# 3. Data

### Libraries

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(leaflet)
library(sf)
library(tigris)
library(zoo)
library(aTSA)
library(tseries)
library(forecast)
```

Get the latest Fairfax calls for service data:
```{r}
calls.full <- read.csv(sprintf("https://docs.google.com/uc?id=%s&export=download",
                               "1ti7BCMe1vvDVVr75p6lqvgouCy_ae7rn"))
```

Format the date and add month columns:
```{r}
calls.full$DATE <- as.Date(calls.full$date, format = "%Y-%m-%d")
calls.full$MONTHS <- months(calls.full$DATE)
calls.full$MONTH <- month(calls.full$DATE)
```

Determine the total count and percentage of each call type:
```{r}
sum.calls <- calls.full %>%
  group_by(type) %>%
  summarise(call.count = n()) %>%
  mutate(pct = round(call.count/sum(call.count)*100,2))
```

Identify a set of calls to analyze, and then filter the data:
```{r}
group <- c("ANIMAL BITE", "ANIMAL CASE", "ANIMAL CONTROL", "ANIMAL LOST-FOUND")
calls.subset <- calls.full %>% filter(calls.full$type %in% group)
```

Calculate calls per day:
```{r}
calls <- calls.subset %>%
  group_by(DATE) %>%
  summarise(CALL.COUNT = n())
```

Fill in blank days:
```{r}
calls <- calls %>%
  complete(DATE = seq.Date(min(DATE), max(DATE), by = "day")) %>%
  mutate(WEEKDAY = lubridate::wday(DATE, label = T, week_start = 1), 
         MONTH = lubridate::month(DATE, label = T, abbr = F),
         WEEK = isoweek(DATE),
         DAY = day(DATE),
         YEAR = year(DATE))
calls <- replace(calls, is.na(calls), 0)
```

Make the data a time series:
```{r}
cleaned.calls <- zoo(calls$CALL.COUNT, 
                       seq(from = as.Date(min(calls$DATE)), 
                           to = as.Date(max(calls$DATE)), by = 1))
```

Generate basic stats and a basic graph:
```{r}
summary(cleaned.calls)
plot(cleaned.calls)
title("Animal Calls Per Day")
```

Make the data stationary:
```{r}
stationary1 <- diff(cleaned.calls, differences = 1)
stationary2 <- diff(cleaned.calls, differences = 2)
stationary3 <- diff(cleaned.calls, differences = 3)
```

Plot each one:
```{r}
plot(stationary1)
plot(stationary2)
plot(stationary3)
```

Run ADF tests on the original data, and the stationary datasets:
```{r}
adf.test(as.matrix(cleaned.calls))
adf.test(as.matrix(stationary1))
adf.test(as.matrix(stationary2))
adf.test(as.matrix(stationary3))
```

Stationary1 is the winner.

Determine the p value using PACF:
```{r}
pacf(stationary1)
pacf(stationary1, pl=FALSE)
```

The p value for this example will be 13.

Determine the q value using ACF:
```{r}
acf(stationary1)
acf(stationary1, pl=FALSE)
```

The q value for this example will be 1.

Build the ARIMA model:
```{r}
arima.calls <- auto.arima(cleaned.calls, d = 1, max.p = 13, max.q = 1, seasonal = T)
summary(arima.calls)
```

Check the model residuals:
```{r}
checkresiduals(arima.calls)
```

Build a sample forecast for 11/1-11/7:
```{r}
forecast.7days <- forecast(arima.calls, h=7)
round(sum(forecast.7days$upper[,2]),0)
forecast.7days$mean
```

Identify our forecast window:
```{r}
forecast.window <- as.numeric(as.Date("2025-12-31")-max(calls$DATE))
```

Forecast 2025
```{r}
forecast.2025 <- forecast(arima.calls, h=forecast.window)
autoplot(forecast.2025)
```

