---
title: "CRIM 515 Final Project Template"
author: "My Name"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  cleanrmd::html_document_clean:
    theme: markdown-modest
    toc: yes
    toc_float: 
      collapsed: true
---

The Final Project follows the same general flow as the Research Projects, as outlined below. Use this format as much or as little as you'd like, as long as the required components are covered.

# 1. Research Question

The goal of this project is to **forecast** future **calls for service** activity in the **City of Fairfax, Virginia** for calendar year 2025. 

These instructions will demonstrate how to forecast using an ***ARIMA model***. An ARIMA model is an Autoregressive Integrated Moving Average, which is a univariate model predicting future values of a single variable over time. Let's breakdown these terms:

  - **Autoregressive**: "auto" means self, and "regressive" refers to using prior values.
  - **Integrated**: "integrated" refers to the process of making data stationary, which identifies trends by differencing the data
  - **Moving Average**: "moving average" is a calculation for a series of averages using subsets of data to identify trends over time. It is a more precise measure for smoothing out fluctuations.

Your research question should consider **behavior**, **space**, and **time**. You'll either *measure* or *control* for each of these. 

- Behavior: what call type(s)
- Space: where
- Time: when is the data from, when are you forecasting

For example, ***using 2008-2024 data, forecast trespass calls for service counts and locations in the City of Fairfax, VA for each month in 2025.***

# 2. Data

### Acquire

1. Load the libraries you need for success:
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(leaflet)
library(sf)
library(tigris)
library(zoo)
library(aTSA)
library(tseries)
library(forecast)
```

2. Get the latest and greatest Fairfax calls for service data:
```{r}
calls.full <- read.csv(sprintf("https://docs.google.com/uc?id=%s&export=download",
                               "1ti7BCMe1vvDVVr75p6lqvgouCy_ae7rn"))
```

### Wrangle

3. Format the date, as a date, and add month columns (different styles) for later:
```{r}
calls.full$DATE <- as.Date(calls.full$date, format = "%Y-%m-%d")
calls.full$MONTHS <- months(calls.full$DATE)
calls.full$MONTH <- month(calls.full$DATE)
```

4. Determine the total count and percentage of each call type:
```{r}
sum.calls <- calls.full %>%
  group_by(type) %>%
  summarise(call.count = n()) %>%
  mutate(pct = round(call.count/sum(call.count)*100,2))
```

5. Use the table from #4 to find and filter the call types to forecast (choose your own!):
```{r}
group <- c("SUSPICIOUS", "MISC")
calls.subset <- calls.full %>% filter(calls.full$type %in% group)
```

# 3. Methods

Building a relevant methodology is largely dependent on your research question and data. Any good methodology involves multiple steps for **acquiring** (done above), **wrangling** (done above), **calculating**, **visualizing**, and **analyzing** data. There is no single right answer here; rather, its the ability to craft a unique, distinct workflow that results in innovative work. 

Some important considerations for ***how*** you use your data include:

  - **SPACE**:
      - Overall: ***Fairfax, VA***
      - Unit of Measurement: ***Neighborhoods? Streets?***
  
  - **TIME**: 
      - Overall: time range for your datas
      - Unit of Measurement: ***Days, Months, Years***
  
  - **BEHAVIOR**:
      - Overall: Types, categories
      - Unit of Measurement: ***Call type(s)***

To use an ARIMA model, we will identify values for **p**, **d**, and **q**:

  - *p*: Lagged observations
  - *d*: Differencing the data set
  - *q*: Prior error handling 

### Calculate

1. Calculate calls per day:
```{r}
calls <- calls.subset %>%
  group_by(DATE) %>%
  summarise(CALL.COUNT = n())
```

2. Fill in blank days:
```{r}
calls <- calls %>%
  complete(DATE = seq.Date(min(DATE), max(DATE), by = "day")) %>%
  mutate(WEEKDAY = lubridate::wday(DATE, label = T, week_start = 1), 
         MONTH = lubridate::month(DATE, label = T, abbr = F),
         WEEK = isoweek(DATE),
         DAY = day(DATE),
         YEAR = year(DATE))
# replace the NAs with 0s
calls <- replace(calls, is.na(calls), 0)
```

3. Change the data type to a proper time series, update the sequencing:
```{r}
cleaned.calls <- zoo(calls$CALL.COUNT, 
                       seq(from = as.Date(min(calls$DATE)), 
                           to = as.Date(max(calls$DATE)), by = 1))
```

4. Generate basic summary stats and a basic graph:
```{r}
summary(cleaned.calls)
plot(cleaned.calls)
title("My Calls Per Day") # this adds a title to your graph
```

5. Make the data stationary (normalized, as a deviation from previous values), and a new basic graph. You can do more/higher differences as necessary (but probably don't need to). Examine the graphs and compare the spikes. These are just examples:
```{r}
stationary1 <- diff(cleaned.calls, differences = 1)
plot(stationary1)
stationary2 <- diff(cleaned.calls, differences = 2)
plot(stationary2)
stationary3 <- diff(cleaned.calls, differences = 3)
plot(stationary3)
```

Look for the data with the narrowest y axis range, as consistently as close to 0 as possible. In this example, ***stationary1*** is the best dataset.

6. Conduct an [Augmented Dickey-Fuller test](https://en.wikipedia.org/wiki/Augmented_Dickey%E2%80%93Fuller_test) to determine if the data is stationary. The resultant score should be a negative number, and the lower the number the stronger the data is stationary. In this example, 'cleaned.calls' is run as a comparison to data that is not differenced.
```{r}
adf.test(as.matrix(cleaned.calls))
adf.test(as.matrix(stationary1))
adf.test(as.matrix(stationary2))
adf.test(as.matrix(stationary3))
```

In this example, all scenarios are negative. When combined with the plot analysis in Step 5, ***stationary1*** is best. Ideally you want to manipulate the data as little as possible. While 'stationary2' and 'stationary3' are even more negative, they are not negative *enough* to warrant choosing them.

7. Lagged autocorrelations - specifically using the Autocorrelation (ACF) and Partial Autocorrelation (PACF) functions - will help determine the **p** and **q** values. For both calculations and visuals, you will use the best stationary dataset you chose in Steps 5 and 6 above.

  - Determine the **p** order by examining the PACF graph and data. Specifically, look for the last *significant* lag:
```{r}
pacf(stationary1)
pacf(stationary1, pl=FALSE)
```

  - Determine the **q** order by identifying the last *significant* lag in the ACF:
```{r}
acf(stationary1)
acf(stationary1, pl=FALSE)
```

In this example, the **p** should be **8**, and the **q** should be **1**.

8. Build an ARIMA function using the inputs derived from previous steps (based on analysis from the graph in Step 4, it's reasonable to assume our data has season trends - so we'll set it to TRUE):
```{r}
arima.calls <- auto.arima(cleaned.calls, d = 1, max.p = 8, max.q = 1, seasonal = T)
summary(arima.calls)
```

Note that the 'auto.arima' function runs a series of ARIMA models and choose the best fit for the data. The best fitting model can be interpreted as ARIMA(d, p, q).

9. Do a quick residuals checks on the model to see how it performed:
```{r}
checkresiduals(arima.calls)
```
Proper residuals derived from your model include:

  - the top graph not demonstrating any clear, cyclical, repeating patterns
  - the bottom right graph generally being a normally distributed bell curve

8. Use the ARIMA function to do a sample forecast for the next week of data:
```{r}
# h = the number of units of time to measure
forecast.7days <- forecast(arima.calls, h=7)
round(sum(forecast.7days$upper[,2]),0)
forecast.7days$mean
```

9. Identify the number of days between the last date in your dataset and the end of 2025:
```{r}
forecast.window <- as.numeric(as.Date("2025-12-31")-max(calls$DATE))
```

10. Forecast the number of calls per day for everyday in that window, and plot it:
```{r}
forecast.2025 <- forecast(arima.calls, h=forecast.window)
autoplot(forecast.2025)
```

11. Extract the forecasted values as a table, clean up the column names, add the forecast date, and month:
```{r}
forecast.values <- as.data.frame(forecast.2025$mean)
forecast.values$ID <- seq.int(nrow(forecast.values))
forecast.upper <- as.data.frame(forecast.2025$upper)
forecast.upper$ID <- seq.int(nrow(forecast.upper))
forecast.values <- forecast.values %>%
  left_join(forecast.upper, by = 'ID')
colnames(forecast.values) <- c("MEAN", "ID", "CI80", "CI90")
forecast.values$DATE <- as.Date(max(calls$DATE) + forecast.values$ID)
forecast.values$MONTH <- months(forecast.values$DATE)
```

12. Filter to 2025, summarize forecasts by month:
```{r}
forecast.values.2025 <- subset(forecast.values, forecast.values$DATE > '2024-12-31')
forecast.months <- forecast.values.2025 %>%
  group_by(MONTH) %>%
  summarise(MEAN = round(sum(MEAN),0), FORECAST.90 = round(sum(CI90),0), FORECAST.80 = round(sum(CI80),0))
forecast.months$DIFF <- forecast.months$FORECAST.90 - forecast.months$FORECAST.80
```

### Visualize

1. Graph of calls with a trend line:
```{r, message=FALSE}
graph.calls <- ggplot(calls, aes(x=DATE, y=CALL.COUNT)) + 
  geom_line() + 
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") + 
  xlab("Years") + 
  ylab("Call Count") + 
  ggtitle("TITLE HERE") + 
  geom_area(fill="lightblue", color="black")

graph.calls + 
  geom_smooth(method = lm, col = "red", se = FALSE) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

2. Graph actual calls with predicted calls:
```{r}
calls2024 <- calls[c(1,2)]
calls2025 <- forecast.values[c(5,1)]
names(calls2025) <- c("DATE", "CALL.COUNT")

new.calls <- rbind(calls2024, calls2025)

graph.new.calls <- ggplot(new.calls, aes(x=DATE, y=CALL.COUNT)) + 
  geom_line() + 
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") + 
  xlab("Years") + 
  ylab("Call Count") + 
  ggtitle("TITLE HERE") + 
  geom_area(fill="lightblue", color="black")

graph.new.calls + 
  geom_smooth(method = lm, col = "red", se = FALSE) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

3. Tweak the data table for making a graph:
```{r}
forecast.months$MONTH <- factor(forecast.months$MONTH, levels = forecast.months$MONTH)
forecast.long <- pivot_longer(forecast.months, cols = c(FORECAST.80, DIFF), 
                          names_to = "Category", values_to = "Value")
forecast.long$Category <- gsub("DIFF", "90% Confidence Interval", forecast.long$Category)
forecast.long$Category <- gsub("FORECAST.80", "80% Confidence Interval", forecast.long$Category)
```

4. Graph your monthly forecasts, first for the upper bounds, and then for the mean:
```{r}
ggplot(forecast.long, aes(x = MONTH, y = Value, fill = fct_rev(Category))) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Value), size = 3, colour = 'white', position = position_stack(vjust = 0.5)) + 
  labs(title = "City of Fairfax 2025 Monthly Forecast Upper Bounds", 
       x = "Month", 
       y = "Call Count") +
  scale_fill_manual(values = c("90% Confidence Interval" = "blue",
                               "80% Confidence Interval" = "grey"),
                    name = "Forecasts") +
  scale_x_discrete(limits = c("January", "February", "March", "April", "May", "June", "July", "August",
                              "September", "October", "November", "December")) + 
  coord_cartesian(ylim = c(0, 500)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplot(forecast.months, aes(x = MONTH, y = MEAN)) +
  geom_bar(stat = "identity") +
  scale_x_discrete(limits = c("January", "February", "March", "April", "May", "June", "July", "August",
                              "September", "October", "November", "December")) +
  coord_cartesian(ylim = c(100, 175)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "City of Fairfax 2025 Mean Monthly Forecast", 
       x = "Month", 
       y = "Call Count")
```

5. Map your monthly hotspots.

  - Get the city outline and roads
```{r, warning=FALSE, message=FALSE, results='hide'}
fairfax.roads <- roads("VA", "Fairfax city")
fairfax.outline <- county_subdivisions("VA", "Fairfax city")
```
  - Filter your data by month
  - Choose a hotspot method
  - Facet by year
```{r, warning=FALSE}
ggplot() +
  geom_sf(data = fairfax.outline, color = "grey") +
  geom_hex(aes(x = lon, y = lat), data = calls.subset, bins = 8) +
  scale_fill_continuous(type = "viridis") +
  geom_sf(data = fairfax.roads, color = "black", alpha = 0.4) +
  theme_classic() +
  facet_wrap(~ MONTH)
```

  - Repeat for multiple years to confirm any monthly patterns.

# 4. Findings

### INSTRUCTIONS

  - Describe the model
  - Describe the outputs
  - Contextualize the results
  - Describe the hotspots
  - Make it real
  - Make it usable

# Writing Tips

Related, some general notes about writing. Welcome to grad school, where the quality of your writing is held to a higher standard. Write directly. Thus, avoid using these words and phrases:

* Simply. Very. Indeed. Specifically. Perhaps. Arguably. Usually. Essentially. Could possibly. Could potentially. Necessarily. Only. Especially. In order to. Likely. Primarily. Generally. It is important that. It is essential to note.

Also, words have meanings. [Some are stronger than others](https://www.themuse.com/advice/185-powerful-verbs-that-will-make-your-resume-awesome). [Others aren’t real](https://www.rd.com/article/words-that-arent-words/).

Further, here’s some other strategies to consider:

* Read sentences aloud… but don’t write like you speak.
* Ask yourself - does this sentence make sense, by itself.
* It’s ok to occasionally use quotes, don’t copy specific language or words.
* If you don’t know the meaning of a word, don’t use it!

And finally, [this is just fun](https://www.thisworddoesnotexist.com/).

# Creating the Output

1. When submitting your projects, you are submitting either an **HTML** or **PDF** output of your Markdown file, as an upload to Canvas - not via email.
2. When you are ready to see what the output looks like, **click the 'Knit' button.** The output file will be saved to the same folder this script is saved to.
3. *Never ever ever include any install.packages functions.* Either comment them out with a #, or delete them.

# Grading: 30pts

#### 1. Research Question
***2pt***

Provide a brief paragraph (1-2 sentences) on the research question. State the question, in terms of behavior, space, and time. Define key terms and concepts, to include how variables will be measured.

#### 2. Literature Review
***2pts***

Provide a brief paragraph describing a prior study related to your current project. This study should be from a peer-academic academic journal. Include details about the *research question*, *data*, *methods*, and *findings*. Cite the article in *APA format*. 

#### 3. Data 
***1pt***

Provide a brief paragraph on the data used in this study. This includes the specific source(s), and the specific temporal and spatial constraints. Address any major advantages and disadvantages with using these data compared to other sources. Be specific enough that a reader could replicate this work.

#### 4. Methods
***2pts***

Provide a brief paragraph on the research methods leveraged to answer this question. This includes the software, calculations, skills, techniques, and unique workflows used to analyze your data and develop an answer. You do NOT need to describe click-by-click instructions or lines of code describing how you did things; you DO need to describe a logical process that is specific enough that a reader could replicate. 

#### 5. Findings
***12pts***

Provide one brief paragraph (4-5 sentences) per month forecasted (*2pts each*). Each paragraph should include forecasted counts, and specific areas of the city to focus on. Your findings should be descriptive, clear, and comprehensively answer the original question. 

#### 7. Visuals
***10pts***

  - Provide a graph that highlights your model (*2pts*).

  - Provide a graph showing the forecasted counts for each month in 2025. This graph should have a title and correct axis labels (*2pts*).

  - Provide hotspot maps (*6pts*). 

  - All visuals should be thoughtful and logical (make sense to someone that has never seen them before); somehow, someway referenced in the text; and analytically meaningful (provide useful, relevant, and actionable insights) 

#### 8. Formatting
***1pt***

- Error-free writing. No typos, run-on sentences, or sentence fragments. Proper punctuation. Real words. Need help paraphrasing dense content? [Try this](https://quillbot.com/). And [here’s a great reference](https://cognella-titles-sneakpreviews.s3-us-west-2.amazonaws.com/83869-1A-URT/83869-1A_SP.pdf), too.
- File submitted as the knitted HTML output of an RMarkdown file, via Canvas.

ONE LAST THING... the source code for creating this file can be [found here](https://github.com/matthew-danna/criminal-justice-research-methods-data-analysis/blob/main/scripts/2024%20fall/CRIM%20515%20Final%20Project%20Template.Rmd).

Please [email me](mailto:mdanna2@gmu.edu) with any questions.
