---
title: "CRIM 515 Final Project In-Class"
author: "Your Name"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  cleanrmd::html_document_clean:
    theme: markdown-modest
    toc: yes
    toc_float: 
      collapsed: true
---

Some R Markdown styles to try, if you're interested:

“almond”, “awsm.css”, “axist”, “bamboo”, “bullframe”, “holiday”, “kacit”, “latex.css”, “markdown-splendor”, “markdown-retro”, “markdown-air”, “markdown-modest”, “marx”, “minicss”, “new.css”, “no-class”, “picocss”, “sakura”, “sakura-vader”, “semantic”, “simplecss”, “style-sans”, “style-serif”, “stylize”, “superstylin”, “tacit”, “vanilla”, “water”, “water-dark”, “writ”

# 1. Research Question

EXAMPLE:  Using 2008-2024 data, this project forecasts animal-related calls for service counts and locations in the City of Fairfax, VA for each month in 2025.

# 2. Literature Review

INSERT THAT TEXT, WITH CITATION, HERE.

# 3. Data

### Libraries

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(leaflet)
library(sf)
library(tigris)
library(zoo)
library(aTSA)
library(tseries)
library(forecast)
```

Get the latest Fairfax calls for service data:
```{r}
calls.full <- read.csv(sprintf("https://docs.google.com/uc?id=%s&export=download",
                               "1ti7BCMe1vvDVVr75p6lqvgouCy_ae7rn"))
```

Format the date and add month columns:
```{r}
calls.full$DATE <- as.Date(calls.full$date, format = "%Y-%m-%d")
calls.full$MONTHS <- months(calls.full$DATE)
calls.full$MONTH <- month(calls.full$DATE)
```

```{r, echo=FALSE, results='hide'}
sum.calls <- calls.full %>%
  group_by(type) %>%
  summarise(call.count = n()) %>%
  mutate(pct = round(call.count/sum(call.count)*100,2))
```

Identify a set of calls to analyze, and then filter the data:
```{r}
group <- c("ANIMAL BITE", "ANIMAL CASE", "ANIMAL CONTROL", "ANIMAL LOST-FOUND")
calls.subset <- calls.full %>% filter(calls.full$type %in% group)
```

Calculate calls per day:
```{r}
calls <- calls.subset %>%
  group_by(DATE) %>%
  summarise(CALL.COUNT = n())
```

Fill in blank days:
```{r}
calls <- calls %>%
  complete(DATE = seq.Date(min(DATE), max(DATE), by = "day")) %>%
  mutate(WEEKDAY = lubridate::wday(DATE, label = T, week_start = 1), 
         MONTH = lubridate::month(DATE, label = T, abbr = F),
         WEEK = isoweek(DATE),
         DAY = day(DATE),
         YEAR = year(DATE))
calls <- replace(calls, is.na(calls), 0)
```

Make the data a time series:
```{r}
cleaned.calls <- zoo(calls$CALL.COUNT, 
                       seq(from = as.Date(min(calls$DATE)), 
                           to = as.Date(max(calls$DATE)), by = 1))
```

Generate basic stats and a basic graph:
```{r}
summary(cleaned.calls)
plot(cleaned.calls)
title("Animal Calls Per Day")
```

Make the data stationary:
```{r}
stationary1 <- diff(cleaned.calls, differences = 1)
```

Plot each one:
```{r}
plot(stationary1)
```

Run ADF tests on the original data, and the stationary datasets:
```{r}
adf.test(as.matrix(stationary1))
```

Stationary1 is the winner.

Determine the p value using PACF:
```{r}
pacf(stationary1)
pacf(stationary1, pl=FALSE)
```

The p value for this example will be 13.

Determine the q value using ACF:
```{r}
acf(stationary1)
acf(stationary1, pl=FALSE)
```

The q value for this example will be 1.

Build the ARIMA model:
```{r}
arima.calls <- auto.arima(cleaned.calls, d = 1, max.p = 13, max.q = 1, seasonal = T)
summary(arima.calls)
```

Check the model residuals (probably not worth keeping in your Final):
```{r, echo=FALSE, results='hide'}
checkresiduals(arima.calls)
```

Identify our forecast window:
```{r}
forecast.window <- as.numeric(as.Date("2025-12-31")-max(calls$DATE))
```

Forecast 2025
```{r}
forecast.2025 <- forecast(arima.calls, h=forecast.window)
autoplot(forecast.2025)
```

Make a forecast table:
```{r}
forecast.values <- as.data.frame(forecast.2025$mean)
forecast.values$ID <- seq.int(nrow(forecast.values))

forecast.upper <- as.data.frame(forecast.2025$upper)
forecast.upper$ID <- seq.int(nrow(forecast.upper))

forecast.values <- forecast.values %>%
  left_join(forecast.upper, by = 'ID')

colnames(forecast.values) <- c("MEAN", "ID", "CI80", "CI95")
forecast.values$DATE <- as.Date(max(calls$DATE) + forecast.values$ID)
forecast.values$MONTH <- months(forecast.values$DATE)
```

Filter to 2025, summarize by month:
```{r}
forecast.values.2025 <- subset(forecast.values, forecast.values$DATE > '2024-12-31')

forecast.months <- forecast.values.2025 %>%
  group_by(MONTH) %>%
  summarise(MEAN = round(sum(MEAN),0), FORECAST.95 = round(sum(CI95),0), FORECAST.80 = round(sum(CI80),0))

forecast.months$DIFF <- forecast.months$FORECAST.95 - forecast.months$FORECAST.80
```

# Graphs

1. Calls per day with a trend line:
```{r}
graph.calls <- ggplot(calls, aes(x=DATE, y=CALL.COUNT)) + 
  geom_line() + 
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") + 
  xlab("Years") + 
  ylab("Call Count") + 
  ggtitle("TITLE HERE") + 
  geom_area(fill="lightblue", color="black")

graph.calls + 
  geom_smooth(method = lm, col = "red", se = FALSE) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

2. Calls per day with a trend line and the forecast:
```{r}
calls2024 <- calls[c(1,2)]
calls2025 <- forecast.values[c(5,1)]
names(calls2025) <- c("DATE", "CALL.COUNT")

new.calls <- rbind(calls2024, calls2025)

graph.new.calls <- ggplot(new.calls, aes(x=DATE, y=CALL.COUNT)) + 
  geom_line() + 
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") + 
  xlab("Years") + 
  ylab("Call Count") + 
  ggtitle("TITLE HERE") + 
  geom_area(fill="lightblue", color="black")

graph.new.calls + 
  geom_smooth(method = lm, col = "red", se = FALSE) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

3. Monthly upper bounds graph:
```{r}
forecast.months$MONTH <- factor(forecast.months$MONTH, levels = forecast.months$MONTH)
forecast.long <- pivot_longer(forecast.months, cols = c(FORECAST.80, DIFF), 
                          names_to = "Category", values_to = "Value")
forecast.long$Category <- gsub("DIFF", "95% Confidence Interval", forecast.long$Category)
forecast.long$Category <- gsub("FORECAST.80", "80% Confidence Interval", forecast.long$Category)

ggplot(forecast.long, aes(x = MONTH, y = Value, fill = fct_rev(Category))) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Value), size = 3, colour = 'white', position = position_stack(vjust = 0.5)) + 
  labs(title = "City of Fairfax 2025 Monthly Forecast Upper Bounds", 
       x = "Month", 
       y = "Call Count") +
  scale_fill_manual(values = c("95% Confidence Interval" = "gold",
                               "80% Confidence Interval" = "maroon"),
                    name = "Forecasts") +
  scale_x_discrete(limits = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")) + 
  coord_cartesian(ylim = c(0, 175)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

4. Monthly forecasts by mean predicted value:
```{r}
ggplot(forecast.months, aes(x = MONTH, y = MEAN)) +
  geom_bar(stat = "identity") +
  scale_x_discrete(limits = c("January", "February", "March", "April", "May", "June", "July", "August",
                              "September", "October", "November", "December")) +
  coord_cartesian(ylim = c(0, 50)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "City of Fairfax 2025 Mean Monthly Forecast", 
       x = "Month", 
       y = "Call Count")
```

Personally, I would use graph #2 AND either graph #3 or graph #4 (so two graphs total).

# Maps

Get the geographic data:
```{r}
fairfax.roads <- roads("VA", "Fairfax city")
fairfax.outline <- county_subdivisions("VA", "Fairfax city")
```

Make the hotspots by month:
```{r}
ggplot() +
  geom_sf(data = fairfax.outline, color = "grey") +
  geom_hex(aes(x = lon, y = lat), data = calls.subset, bins = 6) +
  scale_fill_continuous(type = "viridis") +
  geom_sf(data = fairfax.roads, color = "black", alpha = 0.4) +
  theme_classic() +
  facet_wrap(~ MONTH)
```

Or, make hotspots by month using only "recent" data (where recent is post-COVID or a similar change to your data):
```{r}
ggplot() +
  geom_sf(data = fairfax.outline, color = "grey") +
  geom_hex(aes(x = lon, y = lat), data = subset(calls.subset, calls.subset$year > '2019'), 
           bins = 6) +
  scale_fill_continuous(type = "viridis") +
  geom_sf(data = fairfax.roads, color = "black", alpha = 0.4) +
  theme_classic() +
  facet_wrap(~ MONTH)
```

Don't include both sets of these maps - either use the full time range OR the post-COVID one. Not both.
