---
title: "CRIM 515 Project 2 Template"
author: "My Name"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---
## Libraries
```{r}
library(gtrendsR)
```


## Research Question
Just text, 3 components...

1. how police shootings differ by victims and places over time;

2. if public interest about police shootings is changing; and 

3. describe the relationship between trends in events and public interest in police shootings.

## Literature Review
Just text, one citation, one paragraph

## Data
Some text about acquiring and cleaning three different datasets, and some code

###### Dataset 1
```{r}

```
###### Dataset 2
```{r}

```
###### Dataset 3
```{r}

```

## Methods
Some text about stats and techniques, and some code
```{r}

```

## Findings
#### Finding 1
Some analysis, and some code for Visual 1
```{r}

```

#### Finding 2
Some analysis, and some code for Visual 2
```{r}

```

#### Finding 3
Some analysis, and some code for Visual 3
```{r}

```

## So What
All text, two paragraphs (roughly), all text


# Example Section
```{r, echo=FALSE, message=FALSE, results='hide', fig.show='hide'}
# when you set 'echo' to false, you are hiding the block of code 
# when you set 'message' to false, you are hiding the console text that comes after running your code
# when you set 'results' to hide, you are hiding extra processing that occasionally shows up in the console
# when you set fig.show to hide, you are hiding the plot output
```
